---
title: "Wood density calculations"
author: 'Florian Hartig'
date:  
output:
  html_document: default
---


```{r message =FALSE}
rm(list=ls(all=TRUE))
source("../code/01-packages.R", print.eval =F)
load(file = "../data/cleanedData2.RData")
woodDensity <- read.csv2("../data/GlobalWoodDensityDatabase.csv")
colnames(woodDensity)[which(colnames(woodDensity) == "Wood.density..g.cm.3...oven.dry.mass.fresh.volume")] = "Density"
```


## Deriving wood density values 

The wood density dataset has often several observations per species. We create a summary, using the median per species

```{r}
woodDensitySummarized <- ddply(woodDensity, .(Binomial, Family), summarise,
             medianDensity = median(Density, na.rm = TRUE))
colnames(woodDensitySummarized)[1] <- "Species"
```

Calculate the wood density for each species.

If a species is not in the databse, I use the mean value for the region "South-East Asia (tropical)" (around 0.57). This was the case for approximately 1/3 of the individuals. 


```{r}
speciesData$Species = paste(speciesData$genus, speciesData$specificname)
speciesData$Density = NA
speciesData$DensityDirect = NA

# getting mean wood density 
meanRegion = mean(woodDensity$Density[woodDensity$Region == "South-East Asia (tropical)"])

for (i in 1:length(speciesData$Species)){
  pos = which(as.character(speciesData$Species[i]) == as.character(woodDensitySummarized$Species))
  if (length(pos) == 0){
    speciesData$Density[i] = meanRegion  
    speciesData$DensityDirect[i] = F
  } else if (length(pos) == 1){
    speciesData$Density[i] = woodDensitySummarized$medianDensity[pos]
    speciesData$DensityDirect[i] = T
  } 
}

cleanedData$Density = NA
cleanedData$DensityDirect = NA


for (i in 1:length(cleanedData$Density)){
    pos = which(as.character(cleanedData$Species[i]) == as.character(speciesData$Species))
    if (length(pos) == 0){
    cleanedData$Density[i] = meanRegion  
    cleanedData$DensityDirect[i] = F
    } else if (length(pos) == 1){
    cleanedData$Density[i] = speciesData$Density[pos]    
    cleanedData$DensityDirect[i] = speciesData$DensityDirect[pos] 
  }     
}
```


```{r, fig.width=8, fig.height=8}
par(mfrow=c(2,2))
stages = levels(cleanedData$Stage)
hist(cleanedData$Density, breaks = 50, col = "grey", main = "Frequency of wood density in all plots", xlim = c(0.2,1))
for (i in 1:3)
hist(cleanedData$Density[cleanedData$Stage == stages[i]], breaks = 50, col = "grey", main = paste("Frequency of wood density in", stages[i]), xlim = c(0.2,1))
```


## Form factor 

As a first approximation, I set the form factor to 0.7, assuming a formula of 

AGB = F * BA * Height * Density

We could in the future play around with different form factors. It may make a difference for the estimated absolute biomass values on the plot, but probably not for relative values, and analyses about variability. 

References see 

Chave, J.; Andalo, C.; Brown, S.; Cairns, M.; Chambers, J.; Eamus, D.; Fšlster, H.; Fromard, F.; Higuchi, N.; Kira, T.; Lescure, J.-P.; Nelson, B.; Ogawa, H.; Puig, H.; RiŽra, B. & Yamakura, T. (2005) Tree allometry and improved estimation of carbon stocks and balance in tropical forests. Oecologia, Springer-Verlag, 145, 87-99-.

Brown, S.; Gillespie, A. J. & Lugo, A. E. (1989) Biomass estimation methods for tropical forests with applications to forest inventory data. Forest science, Society of American Foresters, 35, 881-902.


Latifah, Sitti. Inventory and quality assessment of tropical rainforests in the Lore Lindu national park (Sulawesi, Indonesia). Cuvillier Verlag, 2005.

```{r}
cleanedData$FormFactor = 0.7 
```


### Calculating Tree Biomass

```{r}
cleanedData$BasalArea = cleanedData$Dbh_a^2 * pi / 4 * 0.01^2
cleanedData$Biomass = cleanedData$FormFactor * cleanedData$Density * 10^(cleanedData$predictedTreeHeight) * cleanedData$BasalArea 

```


# Analysis


### Basics


```{r}
plotSize = 0.8 * 0.6
ddply(cleanedData, .(Stage, Plot), 
      summarize, 
      Biomass = sum(Biomass) / plotSize ,
      BasalArea = sum(BasalArea) / plotSize )
```

Values should be in t/ha.

The values are relatively low. I think that this is in part due to the relatively low specific wood density of on average around 0.55, which surprised me, but seems to be normal for tropical Asian forests. 


### Variability acros subquadrats for each Plot


```{r, fig.width=8, fig.height=15}
dat <-   ddply(cleanedData, .(Stage, Plot, SubQuad), 
      summarize, 
      Biomass = sum(Biomass) / plotSize * 16 ,
      BasalArea = sum(BasalArea) / plotSize * 16 )

par(mfrow=c(4,2))
selection = levels(cleanedData$Plot)
for (i in 1:length(selection)){
    hist(dat$Biomass[dat$Plot == selection[i]], breaks = 30, col = "grey", main = paste("Biomass variability across subplots in", selection[i]), xlim = c(0,500))
  
}
```


We see increasing variability for the older-growth plots.

I think you can savely conclude that we would need larger inventory plots to get stable biomass estimates. 



# Spatial analysis 


```{r, fig.width=10, fig.height=5}
par(mfrow = c(1,2), mar = c(1,1,1,1))
plots = levels(cleanedData$Plot)
for (i in 1:length(plots)){
  
  selection <- cleanedData[cleanedData$Plot == plots[i],]
  
  maxX = max(selection$X)
  maxY = max(selection$Y)
  
  pattern <- ppp(selection$X, selection$Y, window=owin(c(0,maxX),c(0,maxY)),  marks = selection$Biomass)
  plot(pattern, main = "Biomass Pattern")
  plot(density(pattern, sigma = 1),  main = "Biomass Density")
  mtext(plots[i], outer = T, line = -3, padj = 1)
}

```

Result is pretty much the same as in the subquadrat analysis. 

