---
title: "Wood density calculations"
author: 'Florian Hartig'
date:  `r as.character(format(Sys.Date(), format="%B %d, %Y"))`
output:
  html_document: default
---


```{r message =FALSE}
rm(list=ls(all=TRUE))
source("../code/01-packages.R", print.eval =F)
load(file = "../data/cleanedData2.RData")
woodDensity <- read.csv2("../data/GlobalWoodDensityDatabase.csv")
colnames(woodDensity)[which(colnames(woodDensity) == "Wood.density..g.cm.3...oven.dry.mass.fresh.volume")] = "Density"
```


## Deriving wood density values 

The wood density dataset has often several observations per species. We create a summary, using the median per species

```{r}
woodDensitySummarized <- ddply(woodDensity, .(Binomial, Family), summarise,
             medianDensity = median(Density, na.rm = TRUE))
colnames(woodDensitySummarized)[1] <- "Species"
```

Some species are available for several regions, I haven't summarized this yet. We will take the region 

```{r}
speciesData$Species = paste(speciesData$genus, speciesData$specificname)
speciesData$Density = NA
speciesData$DensityDirect = NA

# getting mean wood density 
meanRegion = mean(woodDensity$Density[woodDensity$Region == "South-East Asia (tropical)"])

for (i in 1:length(speciesData$Species)){
  pos = which(as.character(speciesData$Species[i]) == as.character(woodDensitySummarized$Species))
  if (length(pos) == 0){
    speciesData$Density[i] = meanRegion  
    speciesData$DensityDirect[i] = F
  } else if (length(pos) == 1){
    speciesData$Density[i] = woodDensitySummarized$medianDensity[pos]
    speciesData$DensityDirect[i] = T
  } 
}

cleanedData$Density = NA
cleanedData$DensityDirect = NA


for (i in 1:length(cleanedData$Density)){
    pos = which(as.character(cleanedData$Species[i]) == as.character(speciesData$Species))
    if (length(pos) == 0){
    cleanedData$Density[i] = meanRegion  
    cleanedData$DensityDirect[i] = F
    } else if (length(pos) == 1){
    cleanedData$Density[i] = speciesData$Density[pos]    
    cleanedData$DensityDirect[i] = speciesData$DensityDirect[pos] 
  }     
}
```


```{r, fig.width=8, fig.height=10}
par(mfrow=c(2,2))
stages = levels(cleanedData$Stage)
hist(cleanedData$Density, breaks = 50, col = "grey", main = "Frequency of wood density in all plots", xlim = c(0.2,1))
for (i in 1:3)
hist(cleanedData$Density[cleanedData$Stage == stages[i]], breaks = 50, col = "grey", main = paste("Frequency of wood density in", stages[i]), xlim = c(0.2,1))
```




## Form factor 

As a first approximation, I set the form factor to 0.7, assuming a formula of 

AGB = F * BA * Height * Density

Chave, J.; Andalo, C.; Brown, S.; Cairns, M.; Chambers, J.; Eamus, D.; Fšlster, H.; Fromard, F.; Higuchi, N.; Kira, T.; Lescure, J.-P.; Nelson, B.; Ogawa, H.; Puig, H.; RiŽra, B. & Yamakura, T. (2005) Tree allometry and improved estimation of carbon stocks and balance in tropical forests. Oecologia, Springer-Verlag, 145, 87-99-.

Brown, S.; Gillespie, A. J. & Lugo, A. E. (1989) Biomass estimation methods for tropical forests with applications to forest inventory data. Forest science, Society of American Foresters, 35, 881-902.


Latifah, Sitti. Inventory and quality assessment of tropical rainforests in the Lore Lindu national park (Sulawesi, Indonesia). Cuvillier Verlag, 2005.


The factor 0.1 is for adjusting that we calculate height in m, and Basal area 

```{r}
cleanedData$FormFactor = 0.7 
```


### Calculating Tree Biomass

```{r}
cleanedData$BasalArea = cleanedData$Dbh_a^2 * pi / 4 * 0.01^2
cleanedData$Biomass = cleanedData$FormFactor * cleanedData$Density * 10^(cleanedData$predictedTreeHeight) * cleanedData$BasalArea 

```



# Analysis


### Basics


```{r}
plotSize = 0.8 * 0.6
ddply(cleanedData, .(Stage, Plot), 
      summarize, 
      Biomass = sum(Biomass) / plotSize ,
      BasalArea = sum(BasalArea) / plotSize )
```


### Variability acros subquadrats for each Plot


```{r, fig.width=8, fig.height=15}
dat <-   ddply(cleanedData, .(Stage, Plot, SubQuad), 
      summarize, 
      Biomass = sum(Biomass) / plotSize * 16 ,
      BasalArea = sum(BasalArea) / plotSize * 16 )

par(mfrow=c(4,2))
selection = levels(cleanedData$Plot)
for (i in 1:length(selection)){
    hist(dat$Biomass[dat$Plot == selection[i]], breaks = 30, col = "grey", main = paste("Biomass variability across subplots in", selection[i]), xlim = c(0,500))
  
}
```



# Spatial analysis 


```{r, fig.width=10, fig.height=10}
par(mfrow = c(1,2))
plots = levels(cleanedData$Plot)
for (i in 1:length(plots)){
  
  selection <- cleanedData[cleanedData$Plot == plots[i],]
  
  maxX = max(selection$X)
  maxY = max(selection$Y)
  
  pattern <- ppp(selection$X, selection$Y, window=owin(c(0,maxX),c(0,maxY)),  marks = selection$Biomass)
  plot(pattern, main = "Biomass Pattern")
  plot(density(pattern, sigma = 2),  main = "Biomass Density")
}


```


